ANOVA Comparison Simulations

### 2016-11-29 simulations/simulation.Rmd
Add a few more simulation scenarios

# Test level: simulation under the null

@@

plot_pvalue_hist(design0_pvalues, "No Effect, Normal Errors")

#plot_pvalue_scatter(design0_pvalues, "Simulation P-values for Data with \n No Effect, Normal Errors")
```

# Design 1: Constant additive effect, normal errors

@@

plot_pvalue_hist(design1_pvalues, "Constant Additive Effect of 1, Normal Errors")

#plot_pvalue_scatter(design1_pvalues, "Simulation P-values for Data with \n Constant Additive Effect of 1, Normal Errors")
```

# Design 2: Single site effect, normal errors

@@

plot_pvalue_hist(design2_pvalues, "Effect of 1, Normal Errors")

#plot_pvalue_scatter(design2_pvalues, "Simulation P-values for Data with \n Effect of 1 at a Single Site, Normal Errors")
```

# Design 3: Constant additive effect, heavy-tailed errors

@@

plot_pvalue_hist(design3_pvalues, "Constant Additive Effect of 1, Heavy-tailed Errors")

#plot_pvalue_scatter(design3_pvalues, "Simulation P-values for Data with \n Constant Additive Effect of 1, Heavy-tailed Errors")
```

# Design 4: single site effect, heavy-tailed errors

@@

plot_pvalue_hist(design4_pvalues, "Effect of 1, Heavy-tailed Errors")

#plot_pvalue_scatter(design4_pvalues, "Simulation P-values for Data with \n Effect of 1 at a Single Site, Heavy-tailed Errors")
```

# Design 5: Unpredictive baseline

@@

plot_pvalue_hist(design5_pvalues, "Constant Additive Effect of 1, Normal Errors \n Unpredictive baseline")

#plot_pvalue_scatter(design5_pvalues, "Simulation P-values for Data with \n Constant Additive Effect of 1, Normal Errors")
```

>>>>

## Test level: simulation under the null

@@

plot_pvalue_hist(design0_pvalues, "No Effect, Normal Errors")
```

## Design 1: Constant additive effect, normal errors

@@

plot_pvalue_hist(design1_pvalues, "Constant Additive Effect of 1, Normal Errors")
```

## Design 2: Single site effect, normal errors

@@

plot_pvalue_hist(design2_pvalues, "Effect of 1, Normal Errors")
```

## Design 3: Constant additive effect, heavy-tailed errors

@@

plot_pvalue_hist(design3_pvalues, "Constant Additive Effect of 1, Heavy-tailed Errors")
```

## Design 4: single site effect, heavy-tailed errors

@@

plot_pvalue_hist(design4_pvalues, "Effect of 1, Heavy-tailed Errors")
```

## Design 5: Unpredictive baseline

@@

plot_pvalue_hist(design5_pvalues, "Constant Additive Effect of 1, Normal Errors \n Unpredictive baseline")
```

# Imbalanced designs

Suppose that instead of having 16 individuals per site, we distribute them unequally across sites:
site 1 has 8 patients, site 2 has 16 patients, and site 3 has 24 patients.
This imbalance does not violate any assumptions of the ANOVA.
However, it may reduce power if the effect is concentrated in sites with fewer patients.

## Design 6: Constant additive effect, normal errors

The power curves here are very similar to those in Design 1.
This result is expected, since the effect is still present among all patients who received treatment.

```{r design6, fig.height = 4}
set.seed(760682460) # Generated from random.org Timestamp: 2016-11-14 10:21:12 UTC

tmp <- generate_simulated_data(gamma = 1, effect = "same effect", errors = "normal", n = c(8, 16, 24))
design6_pvalues <- replicate(100, {
  tmp$epsilon <- rnorm(nrow(tmp))
  tmp$Z <- permute_within_groups(tmp$Z, tmp$SITEID)
  tmp$Y1 <- tmp$Y0 + tmp$site_effect + tmp$gamma_vec*tmp$Z + tmp$epsilon
  generate_simulated_pvalues(tmp)
})

design6_pvalues <- t(design6_pvalues)
colnames(design6_pvalues) <- c("ANOVA", "Stratified Permutation", "Differenced Permutation", "LM Permutation", "Freedman-Lane")
design6_power <- apply(design6_pvalues, 2, compute_power)
```

```{r design6_plot, fig.height = 4, fig.align="center"}
plot_power_curves(design6_power, "Constant Additive Effect of 1, Normal Errors")

plot_pvalue_hist(design6_pvalues, "Constant Additive Effect of 1, Normal Errors")
```

## Design 7: Single site effect, normal errors

There is very low power to detect the effect here, since in the simulated data, the site with only 8 patients was the only site with a nonzero effect of treatment.
The treatment effect is on the same order of magnitude as the error variance.

```{r design7, fig.height = 4}
set.seed(760682460) # Generated from random.org Timestamp: 2016-11-14 10:21:12 UTC

tmp <- generate_simulated_data(gamma = 1, effect = "single site effect", errors = "normal", n = c(8, 16, 24))
design7_pvalues <- replicate(100, {
  tmp$epsilon <- rnorm(nrow(tmp))
  tmp$Z <- permute_within_groups(tmp$Z, tmp$SITEID)
  tmp$Y1 <- tmp$Y0 + tmp$site_effect + tmp$gamma_vec*tmp$Z + tmp$epsilon
  generate_simulated_pvalues(tmp)
})
design7_pvalues <- t(design7_pvalues)
colnames(design7_pvalues) <- c("ANOVA", "Stratified Permutation", "Differenced Permutation", "LM Permutation", "Freedman-Lane")
design7_power <- apply(design7_pvalues, 2, compute_power)
```

```{r design7_plot, fig.height = 4, fig.align="center"}
plot_power_curves(design7_power, "Effect of 1 at a Single Site, Normal Errors")

plot_pvalue_hist(design7_pvalues, "Effect of 1, Normal Errors")
```

# Heteroskedastic errors

One assumption of ANOVA is that errors are homoskedastic: they have the same variance across groups.
Suppose that this does not hold in our data.
Imagine we have three sites with 16 patients each, but they have different error distributions:
site 1 has standard normal errors,
site 2 has normally distributed errors with mean 0 and variance 2,
and site 3 has errors with a $t$-distribution on 2 degrees of freedom.

## Design 8: heteroskedastic errors, constant treatment effect

```{r design8, fig.height = 4}
set.seed(760682460) # Generated from random.org Timestamp: 2016-11-14 10:21:12 UTC
nn <- 16

tmp <- generate_simulated_data(gamma = 1, effect = "same effect", errors = "normal", n = rep(nn, 3))
design8_pvalues <- replicate(100, {
  tmp$epsilon[tmp$SITEID == 1] <- rnorm(nn)
  tmp$epsilon[tmp$SITEID == 2] <- rnorm(nn, sd = sqrt(2))
  tmp$epsilon[tmp$SITEID == 3] <- rt(nn, df = 2)
  tmp$Z <- permute_within_groups(tmp$Z, tmp$SITEID)
  tmp$Y1 <- tmp$Y0 + tmp$site_effect + tmp$gamma_vec*tmp$Z + tmp$epsilon
  generate_simulated_pvalues(tmp)
})

design8_pvalues <- t(design8_pvalues)
colnames(design8_pvalues) <- c("ANOVA", "Stratified Permutation", "Differenced Permutation", "LM Permutation", "Freedman-Lane")
design8_power <- apply(design8_pvalues, 2, compute_power)
```

```{r design8_plot, fig.height = 4, fig.align="center"}
plot_power_curves(design8_power, "Constant Additive Effect of 1, Heteroskedastic Errors")

plot_pvalue_hist(design8_pvalues, "Constant Additive Effect of 1, Heteroskedastic Errors")
```

## Design 9: heteroskedastic errors, single site effect

```{r design9, fig.height = 4}
set.seed(760682460) # Generated from random.org Timestamp: 2016-11-14 10:21:12 UTC
nn <- 16

tmp <- generate_simulated_data(gamma = 1, effect = "single site effect", errors = "normal", n = rep(nn, 3))
design9_pvalues <- replicate(100, {
  tmp$epsilon[tmp$SITEID == 1] <- rnorm(nn)
  tmp$epsilon[tmp$SITEID == 2] <- rnorm(nn, sd = sqrt(2))
  tmp$epsilon[tmp$SITEID == 3] <- rt(nn, df = 2)
  tmp$Z <- permute_within_groups(tmp$Z, tmp$SITEID)
  tmp$Y1 <- tmp$Y0 + tmp$site_effect + tmp$gamma_vec*tmp$Z + tmp$epsilon
  generate_simulated_pvalues(tmp)
})
design9_pvalues <- t(design9_pvalues)
colnames(design9_pvalues) <- c("ANOVA", "Stratified Permutation", "Differenced Permutation", "LM Permutation", "Freedman-Lane")
design9_power <- apply(design9_pvalues, 2, compute_power)
```

```{r design9_plot, fig.height = 4, fig.align="center"}
plot_power_curves(design9_power, "Effect of 1 at a Single Site, Heteroskedastic Errors")

plot_pvalue_hist(design9_pvalues, "Effect of 1, Heteroskedastic Errors")
```
